---
title: "NB02: Data Analysis and Visualisation"
output: html_notebook
---

## Confirmatory Analysis: Political Knowledge

```{r}
# Installing necessary packages - remove # if needed/running for the first time
#install.packages("tidyverse")
#install.packages("metafor")
#install.packages("broom")
#install.packages("ggplot2")
```

```{r}
# Loading the packages
library(tidyverse)
library(metafor)
library(broom)
library(ggplot2)
```


```{r}
# Loading the dataset
data_party_cues_IRP <- readRDS("~/Desktop/LSE/Year 3/Independent_Research_Project/Data.nosync/data_party_cues.rds")
```

```{r}
# Filtering for participants who finished the study and completed the captcha to some capacity
df_clean <- data_party_cues_IRP %>%
  filter(Finished == 1, 
         word_captcha %in% c("hello", "Hello", "ehllo", "Ehllo"))
```

```{r}
# Creating a pk_high variable
# Calculating the median of pk_sum, converting it into binary form, saving it as the pk_high variable
df_clean <- df_clean %>%
  mutate(pk_high = ifelse(pk_sum >= median(pk_sum, na.rm = TRUE), 1, 0))
```

```{r}
# Estimating cue treatment effects by political knowledge and issue, fit linear models
effect_sizes <- df_clean %>%
  group_by(item, pk_high) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(outcome_recode_01 ~ cue, data = .x)),
         tidied = map(model, tidy)) %>%
  unnest(tidied) %>%
  filter(term == "cue") %>%
  select(item, pk_high, estimate, std.error)
```

```{r}
# Preparing data for meta-analysis
dat_meta <- effect_sizes %>%
  rename(yi = estimate,
         sei = std.error)
```

```{r}
# Viewing the final dataframe
glimpse(dat_meta)
```


```{r}
# Running the location-scale meta-analytic model
res_mv <- rma.mv(
  yi = yi,                 
  V = sei^2,                   
  mods = ~ pk_high,        
  random = ~ 1 | item,       
  data = dat_meta,
  method = "REML"              
)
```

```{r}
# Summarizing the results
summary(res_mv)
```

```{r}
# Extracting between-issue variance and its SD (heterogeneity estimate)
res_mv$sigma2  
sqrt(res_mv$sigma2)  
```


```{r}
# Adding confidence intervals and readable labels to each estimate
dat_meta <- dat_meta %>%
  mutate(
    ci.lb = yi - 1.96 * sei,
    ci.ub = yi + 1.96 * sei,
    label = paste0("Issue ", item, " (", ifelse(pk_high == 1, "High PK", "Low PK"), ")")
  )
```

```{r}
# Plotting the results for political knowledge
ggplot(dat_meta, aes(x = yi, y = reorder(item, yi))) +
  geom_point(size = 2.5, color = "steelblue3") +
  geom_errorbarh(aes(xmin = ci.lb, xmax = ci.ub), height = 0.2, color = "steelblue3") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "tomato") +
  facet_wrap(~ pk_high, labeller = labeller(pk_high = c("0" = "Low Political Knowledge", "1" = "High Political Knowledge"))) +
  labs(
    title = "Cue Effects by Issue and Political Knowledge",
    subtitle = "Each row is a policy item; 95% confidence intervals shown",
    x = "Effect Size",
    y = "Issue"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    strip.text = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 6),
    panel.spacing = unit(1.5, "lines")
  )
```

```{r}
# Saving the plot
ggsave("../Plots/cue_effects_by_pk.jpeg", width = 10, height = 6, dpi = 300)
```

```{r}
# Density plot to test if it visualizes anything I didm't clearly visualize through the forest plot
ggplot(dat_meta, aes(x = yi, fill = factor(pk_high))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Cue Effects by Political Knowledge",
       x = "Effect Size", fill = "Political Knowledge") +
  scale_fill_manual(values = c("tomato", "steelblue3")) +
  theme_minimal()
```

The above graph was not included in the final version of the paper, as I concluded it did not add any additional insight to the data analysis. However, I still kept it here as a point of reference and supplementary material. It visualizes the density curves of political knowledge groups. 

```{r}
# Saving the second plot
ggsave("../Plots/cue_effects_distribution_pk.jpeg", width = 10, height = 6, dpi = 300)
```

## Exploratory Analysis: Parisanship

```{r}
# Filtering data to include only participants with valid party_bin (Democrat/Republican)
df_party <- df_clean %>% filter(!is.na(party_bin))
```

```{r}
# Estimating cue effects by party and issue
effect_party <- df_party %>%
  group_by(item, party_bin) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(outcome_recode_01 ~ cue, data = .x)),
         tidied = map(model, tidy)) %>%
  unnest(tidied) %>%
  filter(term == "cue") %>%
  select(item, party_bin, estimate, std.error)
```

```{r}
# Preparing partisan meta-analysis data
dat_party_meta <- effect_party %>%
  rename(yi = estimate, sei = std.error) %>%
  mutate(party_bin = factor(party_bin, levels = c("Democrat", "Republican")))
```

```{r}
# Running the location-scale meta-analytic model
res_party <- rma.mv(
  yi = yi,
  V = sei^2,
  mods = ~ party_bin,
  random = ~ 1 | item,
  data = dat_party_meta,
  method = "REML"
)

# Viewing model summary
summary(res_party)
```

```{r}
# Plotting the results for partisanship
dat_party_meta <- dat_party_meta %>%
  mutate(
    ci.lb = yi - 1.96 * sei,
    ci.ub = yi + 1.96 * sei,
    label = paste0("Issue ", item, " (", party_bin, ")")
  )

ggplot(dat_party_meta, aes(x = yi, y = reorder(item, yi))) +
  geom_point(size = 2.5, color = "steelblue3") +
  geom_errorbarh(aes(xmin = ci.lb, xmax = ci.ub), height = 0.2, color = "steelblue3") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "tomato") +
  facet_wrap(~ party_bin) +
  labs(
    title = "Cue Effects by Issue and Partisan Identity",
    subtitle = "95% Confidence Intervals by Self-Reported Party",
    x = "Effect Size",
    y = "Policy Item"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
# Saving the plot
ggsave("../Plots/cue_effects_by_partisanship.jpeg", width = 10, height = 6, dpi = 300)
```


